---
title: "Crime in Europe in 2015"
output:
  html_document: default
  pdf_document: default
  word_document: default
---
```{r download, include=FALSE}
library(dplyr)
library(reshape2)
library(ggplot2)
library(rgdal) # used to read world map data
library(rgeos) # to fortify without needing gpclib
library(maptools)
library(scales) # for formatting ggplot scales with commas
library(eurostat)
library(corrplot)

crime <- read.csv(file="./EurostatCrime2015.csv", header=TRUE, sep=",")
crime <- crime %>% mutate(Sex.crime = Rape + Sexual.assault + Sexual.violence) %>% select(-Rape, -Sexual.assault, -Sexual.violence)
crime <- crime[complete.cases(crime),]
crime <- crime %>% mutate(rank_by_rank = 0, rank_by_proportion = 0, sum_crimes = 0)
for (col in names(crime)[2:6]){
  crime <- crime %>% arrange(eval(parse(text=col))) %>%
    mutate(rank_by_rank = rank_by_rank + row_number(eval(parse(text=col))),
           rank_by_proportion = rank_by_proportion + eval(parse(text=paste0(col,"/sum(",col,")"))),
           sum_crimes = sum_crimes + eval(parse(text=col)))
  crime[,ncol(crime)+1] <- row_number(eval(parse(text=paste("crime$",col))))
  names(crime)[ncol(crime)] <- paste0("rank_",col)
}
crime <- crime %>% mutate(rank_by_rank = rank_by_rank/5,
                          rank_by_proportion = rank_by_proportion/5)
                          

# MAP 1
crime <- crime %>% dplyr::rename(country = X)
crime$country = as.character(crime$country)
crime <- crime %>% rename(Intentional_Homicide = Intentional.homicide,
                          Sex_Crime = Sex.crime,
                          Crime_Score = rank_by_proportion,
                          Crime_Rank = rank_by_rank,
                          Total_Crimes = sum_crimes)
countries <- crime %>% select(country)
countries <- as.vector(countries$country)

# Data from http://thematicmapping.org/downloads/world_borders.php.
# Direct link: http://thematicmapping.org/downloads/TM_WORLD_BORDERS_SIMPL-0.3.zip
# Unpack and put the files in a dir 'data'
worldMap <- readOGR(dsn="./TM_WORLD_BORDERS_SIMPL-0.3",
                    layer="TM_WORLD_BORDERS_SIMPL-0.3")
worldMap.fort <- fortify(worldMap, region = "ISO3")

#'for (c in countries){
#  print(c==worldMap@data[which(worldMap@data[,"NAME"]==c),"NAME"])
#}
# doesn't have Northern Ireland

id <- worldMap@data$ISO3
name <- worldMap@data$NAME
centroids.df <- as.data.frame(coordinates(worldMap))
names(centroids.df) <- c("Longitude", "Latitude")  #more sensible column names
basic <- data.frame(id = id, country=name, centroids.df)
basic$id = as.character(basic$id)
basic$country = as.character(basic$country)
locations <- basic %>% left_join(crime,by="country")
bounds <- basic %>% inner_join(crime,by="country")
```

### Total Crimes
```{r tot_crimes, echo=FALSE}
# Total Crimes
ggplot(locations, aes(map_id = id)) +
  geom_map(aes(fill = Total_Crimes), colour= "grey", map = worldMap.fort) +
  expand_limits(x = worldMap.fort$long, y = worldMap.fort$lat) +
  scale_fill_gradient(high = "red", low = "white", guide = "colorbar", 
                      labels = comma, name="Crimes") +
  geom_text(aes(label = id, x = Longitude, y = Latitude), 
            data = locations[complete.cases(locations),], size=1.9) + 
  coord_equal(xlim = c(min(bounds$Longitude)-2,max(bounds$Longitude)+2), 
              ylim = c(min(bounds$Latitude)-4,max(bounds$Latitude))+4.5) + 
  labs(x = "Longitude", y = "Latitude", title = "Total Crimes (per 1,000 Inhabitants) in 2015, Split by Country") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        legend.title = element_text(size=10),
        legend.text = element_text(size=8, face = "italic"),
        axis.text.y = element_blank(), axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(), axis.title = element_blank())
```

### Correlation Plot
```{r corr, echo=FALSE}
# Correlation Plot
corrmatrix=cor(crime[2:6], use="complete.obs")
rownames(corrmatrix) <- c("Assault", "Homicide", "Robbery", "Theft", "Sex Crime")
colnames(corrmatrix) <- rownames(corrmatrix)
for (i in 1:length(colnames(corrmatrix))){
  colnames(corrmatrix)[i] = paste0(colnames(corrmatrix)[i],"                          ")
}
corrplot(corrmatrix, method="square", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=1,tl.srt=45,addCoef.col="black", addCoefasPercent = TRUE, p.mat = 1-abs(corrmatrix), 
sig.level=0.50, insig = "blank",mar=c(0,0,1,0), diag=TRUE, cl.ratio = 0.25) 
```

Surprisingly the crimes are not very correlated with each other. The only significant correlations (i.e. with a correlation coefficient less than -0.5 or greater than 0.5) being between Assault & Robbery and Theft & Sex Crime.

```{r eurostat, include=FALSE}
query <- search_eurostat("age",type="table")
query[1:20,1:2]
pop <- get_eurostat(id="tps00010", time_format="num")
pop <- label_eurostat(pop)
pop <- pop %>% filter(time==2015) %>% filter(indic_de=="Proportion of population aged 15-24 years") %>%
  rename(country = geo, fifteen_twentyfour = values) %>% select(-indic_de,-time)
pop$country <- as.character(pop$country)
pop <- pop %>% mutate(country = ifelse(grepl('Germany', country),"Germany",country))

query <- search_eurostat("unemployment",type="table")
query[1:20,1:2]
# total unemployment rate
unemployment <- get_eurostat(id="tps00203", time_format="num")
unemployment <- label_eurostat(unemployment)
unemployment <- unemployment %>% filter(time=="2015") %>%
  filter(unit=="Percentage of total population") %>% select(-age, -sex, -unit, -time) %>%
  rename(unemployment_rate = values, country = geo)
unemployment$country <- as.character(unemployment$country)
unemployment <- unemployment %>% mutate(country = ifelse(grepl('Germany', country),"Germany",country))

query <- search_eurostat("debt",type="table")
query[1:27,1:2] %>% print(n=27)
house_debt <- get_eurostat(id="tec00104", time_format="num")
house_debt <- label_eurostat(house_debt)
house_debt <- house_debt %>% filter(time=="2015") %>%
  select(-unit, -na_item, -sector, -time) %>%
  rename(debt_to_inc = values, country = geo)
house_debt$country <- as.character(house_debt$country)
house_debt <- house_debt %>% mutate(country = ifelse(grepl('Germany', country),"Germany",country))

crime2 <- crime %>% select("country","Assault","Intentional_Homicide",
                 "Robbery","Theft","Sex_Crime")
crime2 <- melt(crime2,id.vars = "country") %>% left_join(unemployment,by="country") %>% 
  left_join(house_debt,by="country") %>% left_join(pop,by="country")
crime_exc_theft <- crime2 %>% filter(variable!="Theft")
theft <- crime2 %>% filter(variable=="Theft")

theft <- theft %>% mutate(fifteen_twentyfour = fifteen_twentyfour/100,
                          debt_to_inc = debt_to_inc/100)
```

### Bubble Plot
```{r bubble plot, echo=FALSE}
ggplot(data = theft) +
  geom_point(mapping = aes(x = debt_to_inc, y = fifteen_twentyfour, size=value, colour=value)) +
  scale_color_gradient2(low = "white", midpoint = 2000, mid = "orange", high = "red",limits=c(0, 4200), breaks=seq(0, 4200, by=1000)) +
  guides(color= guide_legend(), size=guide_legend()) +
  scale_size_continuous(limits=c(0, 4200), breaks=seq(0, 4200, by=1000)) +
  labs(y = "Population Aged 15-24", x = "Debt to Income Ratio of Households", title = "Thefts per 1,000 Inhabitants in 2015") +
  scale_x_continuous(labels=percent) +
  scale_y_continuous(labels=percent) +
  theme_bw()
```

Some other Eurostat data indicates that thefts tend to be more prevalent amongest countries with a higher proportion of young people (15 - 24 year olds), as well as ones with more strenuous household debt burdens. The same is not true for the other crimes.

### Plot For Each Crime Type
```{r breakdown1, echo=FALSE}
# Assaults
locations <- locations[complete.cases(locations),]
ggplot(locations, aes(map_id = id)) +
  geom_map(aes(fill = Assault), colour= "grey", map = worldMap.fort) +
  expand_limits(x = worldMap.fort$long, y = worldMap.fort$lat) +
  scale_fill_gradient(high = "red", low = "white", guide = "colorbar", 
                      labels = comma, name="Assaults") +
  geom_text(aes(label = id, x = Longitude, y = Latitude), 
            data = locations[complete.cases(locations),], size=1.9) + 
  coord_equal(xlim = c(min(bounds$Longitude)-2,max(bounds$Longitude)+2), 
              ylim = c(min(bounds$Latitude)-4,max(bounds$Latitude))+4.5) + 
  labs(x = "Longitude", y = "Latitude", title = "Assaults per 1,000 Inhabitants in 2015, Split by Country") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        legend.title = element_text(size=10),
        legend.text = element_text(size=8, face = "italic"),
        axis.text.y = element_blank(), axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(), axis.title = element_blank())
```

```{r breakdown2, echo=FALSE}
# Intentional Homicide
ggplot(locations, aes(map_id = id)) +
  geom_map(aes(fill = Intentional_Homicide), colour= "grey", map = worldMap.fort) +
  expand_limits(x = worldMap.fort$long, y = worldMap.fort$lat) +
  scale_fill_gradient(high = "red", low = "white", guide = "colorbar", 
                      labels = comma, name="Homicides") +
  geom_text(aes(label = id, x = Longitude, y = Latitude), 
            data = locations[complete.cases(locations),], size=1.9) + 
  coord_equal(xlim = c(min(bounds$Longitude)-2,max(bounds$Longitude)+2), 
              ylim = c(min(bounds$Latitude)-4,max(bounds$Latitude))+4.5) + 
  labs(x = "Longitude", y = "Latitude", title = "Intentional Homicides per 1,000 Inhabitants in 2015, Split by Country") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        legend.title = element_text(size=10),
        legend.text = element_text(size=8, face = "italic"),
        axis.text.y = element_blank(), axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(), axis.title = element_blank())
```

```{r breakdown3, echo=FALSE}
# Robbery
ggplot(locations, aes(map_id = id)) +
  geom_map(aes(fill = Robbery), colour= "grey", map = worldMap.fort) +
  expand_limits(x = worldMap.fort$long, y = worldMap.fort$lat) +
  scale_fill_gradient(high = "red", low = "white", guide = "colorbar", 
                      labels = comma, name="Robberies") +
  geom_text(aes(label = id, x = Longitude, y = Latitude), 
            data = locations[complete.cases(locations),], size=1.9) + 
  coord_equal(xlim = c(min(bounds$Longitude)-2,max(bounds$Longitude)+2), 
              ylim = c(min(bounds$Latitude)-4,max(bounds$Latitude))+4.5) + 
  labs(x = "Longitude", y = "Latitude", title = "Robberies per 1,000 Inhabitants in 2015, Split by Country") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        legend.title = element_text(size=10),
        legend.text = element_text(size=8, face = "italic"),
        axis.text.y = element_blank(), axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(), axis.title = element_blank())
```

```{r breakdown4, echo=FALSE}
# Theft
ggplot(locations, aes(map_id = id)) +
  geom_map(aes(fill = Theft), colour= "grey", map = worldMap.fort) +
  expand_limits(x = worldMap.fort$long, y = worldMap.fort$lat) +
  scale_fill_gradient(high = "red", low = "white", guide = "colorbar", 
                      labels = comma, name="Thefts") +
  geom_text(aes(label = id, x = Longitude, y = Latitude), 
            data = locations[complete.cases(locations),], size=1.9) + 
  coord_equal(xlim = c(min(bounds$Longitude)-2,max(bounds$Longitude)+2), 
              ylim = c(min(bounds$Latitude)-4,max(bounds$Latitude))+4.5) + 
  labs(x = "Longitude", y = "Latitude", title = "Thefts per 1,000 Inhabitants in 2015, Split by Country") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        legend.title = element_text(size=10),
        legend.text = element_text(size=8, face = "italic"),
        axis.text.y = element_blank(), axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(), axis.title = element_blank())
```

```{r breakdown5, echo=FALSE}
# Sex Crime
ggplot(locations, aes(map_id = id)) +
  geom_map(aes(fill = Sex_Crime), colour= "grey", map = worldMap.fort) +
  expand_limits(x = worldMap.fort$long, y = worldMap.fort$lat) +
  scale_fill_gradient(high = "red", low = "white", guide = "colorbar", 
                      labels = comma, name="Cases") +
  geom_text(aes(label = id, x = Longitude, y = Latitude), 
            data = locations[complete.cases(locations),], size=1.9) + 
  coord_equal(xlim = c(min(bounds$Longitude)-2,max(bounds$Longitude)+2), 
              ylim = c(min(bounds$Latitude)-4,max(bounds$Latitude))+4.5) + 
  labs(x = "Longitude", y = "Latitude", title = "Sex Crime Incidents per 1,000 Inhabitants in 2015, Split by Country") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        legend.title = element_text(size=10),
        legend.text = element_text(size=8, face = "italic"),
        axis.text.y = element_blank(), axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(), axis.title = element_blank())
```
